Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-05-02 products/manuscript/Manuscript.qmd
Added to Background Info and proofread manuscript

We will be looking for patterns in contraceptive methods and unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,short acting reversible, and traditional. Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies [@jonathanmarcbearakCountryspecificEstimatesUnintended2022]. According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group. Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study [@rivetContraceptionUseAttitudes2021]. African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies. This could be due to the current legal status of contraception within Latin American countries. Free emergency contraception access exists in this region, but differing regulations limit access. Chile, Colombia, and Ecuador legally recognize access to emergency contraception, while Nicaragua and Bolivia rely on Ministerial protocols. Argentina and Brazil lack legal recognition, but offer protocols and guides. Mexico requires provision of emergency contraception to victims of sexual and domestic violence [@heviaLegalStatusEmergency2012]. Stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging [@kavanaughContraceptiveMethodUse2018].

>>>>

We will be looking for patterns in contraceptive methods and unintended pregnancies across all age groups (15-19, 20-24, 25-35, and 35-49) to investigate the effectiveness of three contraceptive methods: long acting reversible and sterilization,short acting reversible, and traditional. Financial, social, religious, and cultural factors are current challenges against efforts towards reducing unintended pregnancies [@jonathanmarcbearakCountryspecificEstimatesUnintended2022]. According to a cross-sectional survey done in 2017, contraceptive use was prevalent among women who had attained higher education, were in stable partnerships, and identified with the dominant ethnic group. Additionally, factors like previous pregnancies and immigration status did not appear to influence contraceptive use in this study [@rivetContraceptionUseAttitudes2021]. African and Latin American/Caribbean countries have the highest percentages of unintended pregnancies. This could be due to the current legal status of contraception within Latin American countries. All Latin American have ratified international agreements that recognize a woman's right to privacy, health, and control over her reproductive choices. However, the legal status of contraception varies greatly by country. Chile, Colombia, and Ecuador legally recognize access to emergency contraception, while Nicaragua and Bolivia rely on legal protocols. In 2008, the Chilean Constitutional Court argued against the legal principle of “giving preference to an interpretation that favors the right of ‘a person’ to life above any other interpretation that may annul such a right.”[@heviaLegalStatusEmergency2012] This decision permitted the use of emergency contraception. Argentina and Brazil lack legal recognition, but offer protocols and guides as part of the “Obligatory Medical Program”. Mexico requires provision of emergency contraception to victims of sexual and domestic violence. Contrastingly, in 2009, Honduras issued a complete ban on the distribution of emergency contraception. [@heviaLegalStatusEmergency2012]. It is important to note that judges are not responsible for examining scientific evidence before making a legal decision. It has been observed that stable contraceptive use among high-risk women suggests reaching even lower national unintended pregnancy rates might be challenging. [@kavanaughContraceptiveMethodUse2018].
